---
title: "MLE Project"
author: "Yiyi Xu"
date: "9/21/2018"
output:
  html_document: default
  pdf_document: default
---
# 3.3.2 Many Local Maxima 

## Find the log-likelihood and plot

Equation: 

$$L(\theta;x)=\prod_{i=1}^{n}f(x_{i};\theta)=(2\pi)^{-n}\prod_{i=1}^{n}[1-cos(x_i-\theta)]$$
$$l(\theta)=\ln L(\theta;X)=-n\ln(2\pi)+\sum_{i=1}^{n}\ln[1-cos(x_i-\theta)]$$
```{r }
sample <- c(3.91, 4.85, 2.28, 4.06, 3.70, 4.04, 5.46, 3.53, 2.28, 1.96,
       2.53, 3.88, 2.22, 3.47, 4.82, 2.46, 2.99, 2.54, 0.52)
l <- function(x, y){
  -19*log(2*pi)+sum(log(1-(cos(x-y))))
}
theta <- seq(-pi, pi, 0.1)
L <- numeric(0)
for (i in 1: length(theta)){
   L[i] <- l(sample,theta[i])
}
 sp=spline(theta,L,n=1000)
 plot(sp,col="red",type="l",xlim=c(-pi,pi),ylim=c(-80,-25),lwd=2,xlab="theta",ylab="l(theta)",sub="loglikehood function against theta",col.main="green",font.main=2)

```
        
## Find Method of moments estimator of theta
$$E[X|\theta]=\int_0^{2\pi}\frac{x(1-cos(x-\theta))}{2\pi}dx=\frac{1}{2\pi}(2\pi^2+2\pi sin(\theta))=\sin(\theta)+\pi$$
$$\tilde{\theta}_n = \arcsin(\bar{X}_n - \pi)$$

## Find the MLE by Newton-Raphson method

$$l'(\theta)=\sum_{i=1}^n \frac{-sin(x_i-\theta)}{1-cos(x_i-\theta)}  $$
$$l''(\theta)=\sum_{i=1}^n \frac{1}{cos(x_i-\theta)-1}$$
  
```{r}
l1 <- function(sample, theta){
  sum((-sin(sample-theta))/(1 -cos(sample-theta)))   
}
l2 <- function(sample,theta){
  sum((1)/(cos(sample-theta)-1))   
}
Newton.Method <- function(y,f,f1){
   y0 <- y
   for(i in 1:100){
    y1 <- y0 - f(sample,y0)/f1(sample,y0)
    if(abs(y1-y0)<0.0001) 
      break
    y0 <- y1
   }
   return(data.frame(init=y,root=y0,iter=i))
}

Newton.Method(pi - asin(mean(sample) - pi),l1,l2)
Newton.Method(asin(mean(sample) - pi),l1,l2)

```
## solutions 
```{r}
Newton.Method(-2.7,l1,l2)
Newton.Method(2.7,l1,l2)
````

## repoeat 200 times 
```{r}
init <- seq(-pi,pi,length=200)
root <- numeric(length(init))
for(i in 1:length(init)){
  root[i] <- Newton.Method(init[i],l1,l2)[1,2]
}
plot(init,root,main = "Roots VS. Initial Values")
```




# 3.3.3 Modeling beetle data 

## Fit the population growth model to the beetles data using the Gauss-Newton approach, to minimize the sum of squared errors between model predictions and observed counts.  From data frame, assues initial N is 2. 

$$F=\sum _{i=1}^n [N_i-\frac{2K}{2+(K-2)e^{-rt_i}}]^2 $$
```{r}
beetles <- data.frame(
    days    = c(0,  8,  28,  41,  63,  69,   97, 117,  135,  154),
    beetles = c(2, 47, 192, 256, 768, 896, 1120, 896, 1184, 1024))
N0 <- 2   ## initial value by beetles. 

func <- function(x){
   beetles$beetles-(2*x[1])/(2+(x[1]-2)*exp(-x[2]*beetles$days))
}

library(pracma)
gaussNewton(c(100,1),func)
```

##  Show the contour plot of the sum of squared errors

```{r}
func1 <- function(a,b){
   sum((beetles$beetles-(2*a)/(2+(a-2)*exp(-b*beetles$days)))^2)
}
 
library(plotly)
k <- seq(0,1500,0.1)
r <- seq(0,0.5,0.001)
y <-outer(k,r,Vectorize(func1))
contour(k,r,y,main="Contour plot of the sume of Squared Errors")


```



## Find the maximum likelihood estimators & Estimate the variance your parameter estimates.
```{r}
lnday <- log(beetles$beetles)
# by previous result, set:
k1 <- 1049.4072453    
r1 <- 0.1182684

lnfunc <- function(x){
  k <- x[1]
  r <- x[2]
  sigma <- x[3]
  -sum(-(log(2 * pi * (sigma)) / 2 )- (log(beetles$beetles) - log((2*k)/(2 + (k - 2) * exp(-r * beetles$days)))) ^ 2 / (2 * (sigma)))
}

x <- c(k1,r1,1)
lnfunc(x)
MLEfunc <- optim(x, lnfunc, method = "BFGS",hessian = TRUE)
MLEfunc



```










